---
output: 
    github_document:
        df_print: kable
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# tidyinftheo

[![Travis-CI Build Status](https://travis-ci.org/pohlio/tidyinftheo.svg?branch=master)](https://travis-ci.org/pohlio/tidyinftheo)
[![codecov](https://codecov.io/gh/pohlio/tidyinftheo/branch/master/graph/badge.svg)](https://codecov.io/gh/pohlio/tidyinftheo)

You can install tidyinftheo from github with:

```R
devtools::install_github("pohlio/tidyinftheo")
```

## Examples

Calculate (in bits) the Shannon Entropy of the eye color variable in the `starwars` dataset:
```{r}
library(tidyinftheo)
starwars %>% shannon_entropy(eye_color)
```
With the classic `mtcars` dataset, choose some columns to make a matrix of mutual information pairwise comparisons.  In particular, the *cyl*, *vs*, *am*, *gear*, and *carb* columns are all whole numbers indicating they belong to a category.  The other columns are continuous and are better suited to correlation comparisons, unless they're discretized.  Here are the first few rows of **mtcars**:

```{r}
mtcars %>% select(cyl, vs, am, gear, carb) %>% head()
```
And here is our comparison table.  There should be *5-choose-2 = 10* different combinations.  NMI stands for *Normalized Mutual Information*, so the mutual information, normally given in bits, is scaled between 0 and 1:
```{r}
mi_matr <- as_tibble(mtcars) %>% 
    mutate_if(is_double, as.character) %>%
    mutual_info_matrix(cyl, vs, am, gear, carb, normalized=TRUE)
mi_matr
```
The matrix is already in a convenient format to plot:
```{r}
axis_names <- mtcars %>% select(cyl, vs, am, gear, carb) %>% names()
p <- mi_matr %>%
    ggplot(aes(V1, V2)) +
    geom_tile(aes(fill=NMI), color="white") +
    scale_x_discrete(limits=axis_names[1:(length(axis_names)-1)]) +
    scale_y_discrete(limits=rev(axis_names)[1:(length(axis_names)-1)]) +
    theme(axis.text.x=element_text(angle=90, hjust=1, size=12),
          axis.text.y=element_text(size=12)) +
    xlab(NULL) + ylab(NULL) + 
    ggtitle("mtcars mutual information comparisons")
print(p)
```
